### 2025-02-26 data/preparation.Rmd
updating data preparation

---
title: "preparation"
author: "Reilly Amera"
date: "2025-02-26"
output: html_document
---

```{r setup}
library(tidyverse)
```

```{r read-in}
child <- read_csv("child.csv", col_names = TRUE)
parent <- read_csv("parent.csv", col_names = TRUE)
```

```{r cleaning}
# removing unnecessary columns
child <- child %>% 
  select(-c(1:5, 8:18))
child <- child[-c(1, 2),]

parent <- parent %>% 
  select(-c(1:6, 8:18))
parent <- parent[-c(1, 2),]

# replacing one NA (participant forgot ID)
child$PID[is.na(child$PID)] <- 1

# renaming 'finished' columns 
child <- child %>% 
  rename(Cfinished = Finished)

parent <- parent %>% 
  rename(Pfinished = Finished)
```

```{r wrangling}
# merging datasets
youtube <- full_join(parent, child, by = "PID")

# removing instances where either parents or children failed to finish
youtube <- youtube %>% 
  filter(Pfinished == Cfinished)
```