### 2025-03-12 manuscript.qmd
updated qmd

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
    address: "1155 E 60th St."
    city: Chicago
    region: IL
    country: USA
    postal-code: 60637
abstract: "One way in which children learn about group dynamics within their racial world is from patterns in their daily environments. The current study seeks to understand the degree to which YouTube, a popular form of social media, exposes children to negatively valenced racial patterns, and potentially affects their beliefs. Dyads of parents and 8-13 year-old children (N = 18) describe daily YouTube usage, and the effects of children's YouTube usage are assessed. In the present sample, there is no relationship between children's YouTube usage, racial patterns they may observe, and negative racial beliefs."
floatsintext: true
numbered-lines: false
bibliography: [r-references.bib, bibliography.bib]
link-citations: true
draft-date: false
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
format:
  apaquarto-html: default
  apaquarto-docx: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    list-of-tables: true
editor: visual
---

```{r}
#| label: setup
#| include: false

# Necessary libraries
library(svglite)
library(papaja)
library(knitr)
library(broom)
#install.packages(kableExtra)
library(kableExtra)
library(flextable)
library(psych)
library(tidyverse)

# Setting global options
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE)

# Generate or overwrite r-reference files
# papaja::r_refs(file = "r-references.bib")

# read in data
youtube <- read_csv("data/youtube.csv", col_names = TRUE)

```

```{r}
#| label: my-theme

my_theme <- function(){
  theme(
    # Plot background color 
    plot.background = element_rect(fill = NA),
    # Adding border 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    # Panel background
    panel.background = element_rect(fill = NA),
    # Plot labels
    plot.title = element_text(color = "black", family = "serif", face = "bold"),
    plot.subtitle = element_text(color = "black", family = "serif"),
    plot.caption = element_text(color = "black", family = "serif", face = "italic"),
    # Customizing grid lines 
    panel.grid.major.y = element_line(color = "lightgrey", linetype = 1, linewidth = 0.5),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "lightgrey", linetype = 1, linewidth = 0.5),
    panel.grid.minor.x = element_blank(),
    # Customizing facets
    strip.background = element_rect(color = "black", fill = NA),
    strip.text = element_text(color = "black", hjust = 0.5, family = "serif"),
    # Customizing axes
    axis.text.y = element_text(color = "black", family = "serif", face = "bold.italic"),
    axis.text.x = element_text(color = "black", family = "serif", face = "bold.italic"),
    axis.title = element_text(color = "black", family = "serif", face = "bold"),
    axis.ticks = element_line(color = "black"),
    # Customizing legend
    legend.title = element_text(color = "black", family = "serif", face = "italic"), 
    legend.text = element_text(color = "black", family = "serif")
  )
}

```

Is YouTube making kids more racist? Despite sounding like the cultural fear-mongering often associated with screen time and development, such a question refers to real concerns that psychological research has yet to address: what are children really learning from YouTube, particularly with regard to race?

## Literature Review {#sec-lit-review}

Much of the literature on how children learn about their racial world places great emphasis on the transmission of information that is explicitly about race: how parents talk about members of other racial groups, how and whether schools portray the histories of marginalized groups, how peers may communicate stereotypes, etc. [@hugheswatford_2022_racial]. In this view, ethnic-racial socialization, or the ways in which agents transmit messages about ethnicity and race to children, occurs through direct verbal instruction. However, in addition to more explicit transmissions of racial knowledge, children are simultaneously exposed to a barrage of more implicit environmental inputs—who they live near, for instance, or who they see on screens—which may also affect these conceptions [@hughesetal_2023_growing].

Not only does ethnic-racial socialization occur with explicit instruction, but so too do children attend to the racial patterns and organizations embedded into their environments and social settings [@hugheswatford_2022_racial]. Such environmental racial patterns—who tends to hold positions of authority, who tends to receive discipline in school, etc.—constitute racial regularities, or “persistent patterns in the relative social experiences of, or roles occupied by, two or more racial (or racialized groups) within a setting” [@hugheswatford_2022_racial]. In response to such patterns, children engage in racial learning, or the meaning-making of external messaging so as to form their own conceptions of how race works in their social world [@hugheswatford_2022_racial]. As such, when exposed to consistent, repeated regularities, children draw inferences as to why such patterns exist and persist; they may accept that such patterns as necessary features of ‘the way things are,’ or may even create explanations for such patterns (e.g. “White families live in big houses because they work hard” in contrast to “Black families live in big houses because they got lucky”) [@hughesetal_2023_growing; @rizzoetal_2022_beliefs]. Further, this learning may be yet more implicit: when tested in the same environment, young children show levels of implicit biases, unconscious mental associations based on race or other social categories, similar to those of adults [@payneetal_2017_bias] This suggests that features of specific environments such as racial regularities not only affect explicitly articulated beliefs but implicit understandings of one’s social environment. As such, to fully understand how children’s conceptions of race form, we must examine the ways in which environments containing ontologically inaccurate racial regularities affect their racial learning.

One such environmental stream of input regarding race and racial regularities is social media. Children today have unparalleled access to social media platforms: YouTube, TikTok, Twitch, and more. On these sites, the sheer amount of content posted often outpaces enforcement of content guidelines, making it difficult to know exactly what children are seeing or learning from the content they are consuming. Unlike media such as television or movies, social media platforms are populated by both media corporations and individual content creators. YouTube is a social media platform that is particularly popular amongst children; many children use YouTube regularly: one third of parents of children aged 11 and younger surveyed by Pew Research Center indicated that they let their children use YouTube regularly [@smithetal_2018_many]. Further, amongst tweens and teens, video logs and gaming videos—types of content usually generated by influencers, not corporations—are the two most popular genres watched [@rollinsetal_2022_who].

Despite massive popularity, research on the effects of YouTube and other forms of social media on children’s race-related attitudes and behaviors has not been methodologically systematic [@wardbridgewater_2023_media]. Though YouTube has become a regular part of children’s lives, it is still unknown how YouTube may be affecting children’s conceptions of racial categories and associated beliefs about race. However, some research has emerged describing racial regularities on YouTube: in a review conducted by Common Sense Media, in YouTube videos watched by children under the age of 8, prominent characters of color were far more likely to engage in violent behaviors or inappropriate language, and were less likely to appear alongside educational content [@rollinsetal_2022_who]. In another content analysis of the most popular YouTube videos at the time, more than three-fifths of videos in the sample included racial stereotypes, most of which reinforced and perpetuated them [@guoharlow_2014_usergenerated]. Taken together, we see that YouTube contains racial regularities in the form of stereotypes, which children may be accommodating into their beliefs, but there has not yet been any research confirming this connection.

While there is a body of research aimed at understanding how certain aspects of children’s environments contribute to the formation of racial attitudes, there is not as much focus on how YouTube, or other social media platforms, function as sites of racial socialization and learning via racial regularities. Similarly, though there have been some content analyses regarding the racial character of YouTube videos, such studies have yet to be done in conjunction with psychological measures of racial beliefs and attitudes. Due to this lack of research into how social media ecologies contribute to kids’ racial beliefs, I thus ask the question: are children for whom stereotypical social media is a part of their daily environment more likely to develop stereotyping or prejudicial racial beliefs and attitudes?

## Present Study {#sec-present-study}

This study will examine the relationship between YouTube usage and children’s development of racial attitudes, so as to understand whether and the degree to which YouTube may be a site of ethnic-racial socialization where children develop prejudices against those of other races.

I hypothesize that time spent on YouTube and children’s observation of racial regularities should covary; as children spend more time on YouTube, they may observe more negatively valenced racial regularities including people of color [@guoharlow_2014_usergenerated; @rollinsetal_2022_who]. Thus, I hypothesize that if children spend more time on YouTube, they will display greater racial stereotyping and prejudice against outgroups. While some children may not be consciously aware of on-screen racial regularities, these dynamics may still affect their racial attitudes [@hugheswatford_2022_racial]. As such, I hypothesize that the relationship between children’s time spent on YouTube and their exhibited racial stereotyping and prejudice will be mediated by their observations of negative racial regularities.

One additional factor that may moderate the relationship between time spent on YouTube and racial attitudes is whether a child’s favorite YouTuber is of a race other than their own, via the parasocial contact hypothesis. This theory suggests that when individuals have limited real life interpersonal contact with minority groups, the formation of parasocial (one-sided) relationships with outgroup media figures may reduce stereotyping and prejudice [@schiappaetal_2005_parasocial; @wongetal_2022_theorizing]. Much of the parasocial contact hypothesis literature explores the prejudice reduction made possible by parasocial relationships with television characters, yet the growing popularity of social media presents another medium through which parasocial relationships may be formed. If positive parasocial contact hypothesis effects may be observed when individuals form relationships with characters in other media, so too should these effects be observed when individuals have parasocial relationships with outgroup influencers [@banasetal_2020_metaanalysis]. Thus, I pose my third hypothesis: the relationship between the amount of time that children spend on YouTube and their racial stereotyping and prejudice will be moderated by whether they have a parasocial relationship with a favorite YouTuber of a race other than their own. Specifically, if children have a favorite YouTuber of a race other than their own and they have a parasocial relationship with this YouTuber, they will display less racial stereotyping and prejudice.

# Methods {#sec-methods}

## Participants and Procedures {#sec-participants-and-procedures}

This study includes dyads of parents and children between the ages of 8 and 13 [^1]. Though children may express negative racial attitudes prior to this point in development, during early adolescence, children develop the ability to recognize and reflect on the racial patterns and dynamics present in their environments [@hugheswatford_2022_racial]. As such, this age range allows me to observe how the observation racial regularities present on YouTube may affect children’s stereotyping and prejudicial attitudes, and how such a phenomenon develops. Participants were recruited through Qualtrics Panels, allowing me to obtain a representative sample of parents and children; other studies have used Qualtrics Panels with similar methods and populations [@byrdahn_2020_profiles].

[^1]: Pilot data has demonstrated that Qualtrics Panels has not, thus far, been as stringent in participant age verification. As such, data used presently is from a broader age range than will be included in the final sample, after stricter age validation has been included.

Pilot data has been collected in March of 2025 from families recruited via Qualtrics Panels (*N* = `r length(youtube$transaction_id)`). Participating families receive two Qualtrics surveys to be completed asynchronously, and receive compensation (as determined by Qualtrics Panels) after both surveys have been completed. First, parents complete a demographics and YouTube usage survey, then children complete their own survey.

## Measures {#sec-measures}

In the parent survey, participants answer questions related to parent and child demographics (child age, race/ethnicity, parent income and level of education, etc), as well as questions regarding their child’s YouTube usage (for instance, time spent on YouTube, as reported by their child’s device screen time).

In the child survey, children answer questions related to their YouTube usage, such as time spent on YouTube and their favorite genre. Next, they will either answer more questions pertaining to YouTube content, or their general racial attitudes; either block of questions will appear randomly, so as to reduce the influence of question order. In the racial attitudes section, children will answer questions regarding outgroup prejudices, as rated on a Likert scale. They will also complete one friend choice item, as a simple exploratory measure of racial preference. In the YouTube content section, children will be asked to think of their favorite YouTuber as they answer questions related to whether they have a parasocial relationship with that YouTuber, as well as their observations of racial regularities on YouTube.

The particular measures used in the child survey are as follows:

1. Beliefs about groups (adapted from @halperinetal_2012_promoting)
2. Perceived homophily in behavior and appearance (adapted from @mccroskeyetal_1975_development)
3. Parasocial interaction with a YouTuber (adapted from @sungetal_2023_shes)
4. Racial regularities (created based on @rollinsetal_2022_who)
5. Friend choice (adapted from @amemiyabian_2024_why)

All measures used in adult studies have been specifically adapted for child readability. @tbl-reliability reports the reliability of all questionnaires used, within the present sample.

```{r}
#| label: report-reliability-function

# Function which returns Cronbach's alpha
return_alpha <- function(data, cols) {
  # select data 
  selected_data <- data %>% select(all_of(cols))
  # compute alpha
  alpha_result <- psych::alpha(selected_data, check.keys = TRUE)
  # Extract and return the raw alpha value
  return(alpha_result$total$raw_alpha)
}
```

```{r}
#| label: reliability-dataframe

# Creating a list of all measures, grouped by questionnaire
measures <-
  list(
  "Group Malleability" = c("malleable_1", "malleable_2", "malleable_3", "malleable_4"),
  "Homophily (Appearance)" = c("homophily_appear_1", "homophily_appear_2", "homophily_appear_3", "homophily_appear_4", "homophily_appear_5"),
  "Homophily (Appearance)" = c("homophily_behave_1", "homophily_behave_2", "homophily_behave_3", "homophily_behave_4", "homophily_behave_6"),
  "Parasociality" = c("parasocial_1", "parasocial_2", "parasocial_3", "parasocial_4", "parasocial_5"),
  "Racial Regularities" = c("regularities_1", "regularities_2", "regularities_3", "regularities_4", "regularities_5")
)

# Creating data frame with measures and corresponding alpha
reliability <- data.frame(
  Measure = names(measures),
  Alpha = sapply(measures, function(cols) return_alpha(youtube, cols))
)

# Adding column with evaluations of alpha
reliability <- reliability %>%
  mutate(Alpha = round(Alpha, digits = 2)) %>% 
  mutate(
    Evaluation = case_when(
      Alpha >= 0.95 ~ "Excellent",
      Alpha >= 0.9 ~ "Great",
      Alpha >= 0.8 ~ "Good",
      Alpha >= 0.7 ~ "Acceptable",
      Alpha >= 0.6 ~ "Questionable",
      Alpha < 0.6 ~ "Poor"
    )
  )
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "Reliability of Questionnaires"

# Generate table
kable(reliability, 
      align = c("c", "c", "c"),
      # Setting caption here
      caption = "Reliability (assessed through Cronbach's Alpha) of questionnaires used, within sample") %>% 
  kable_styling(position = "center", latex_options = c("striped"))

```

Beliefs about groups, particularly in terms of group malleability, were chosen as the primary outcome measure to assess racial attitudes. Such questions, adapted from @halperinetal_2012_promoting, address whether children have fixed or malleable views regarding outgroups (e.g. when prompted to think about groups with different skin colors than their own, children are asked the extent to which they agree with the statement, "These groups of people can't really change the way they act"). Higher scores in this domain demonstrate more fixed evaluations of outgroup members, suggesting an underlying racial essentialism, or a belief in the intrinsic nature of racial difference [@waxman_2021_racial].

Composite scores have been created for each measure. All analyses have been completed in RStudio [@R-base], primarily within the tidyverse and psych packages (@R-tidyverse; @R-psych).

# Results {#sec-results}

Looking first to the distributions of time spent on YouTube among  participants, just as many children use YouTube every day as (*N* = `r sum(youtube$daily == 'Every Day')`) as those who do not (*N* = `r sum(youtube$daily == 'Not Every Day')`). And, as shown in @fig-time-distribution, children who use YouTube every day appear to use YouTube for more time on an average day than children who do not. On average, parents report that children spend about `r round(mean(youtube$Ptime), digits = 0)` minutes, or `r round(mean(youtube$Ptime)/60, digits = 0)` hour, on YouTube on an average day. Some children, as parents report, watch as much as `r round(max(youtube$Ptime)/60, digits = 0)` hours of YouTube, while others watch only `r round(min(youtube$Ptime)/60, digits = 0)`. Children's self-report YouTube usage reflects a similar pattern, as shown in @tbl-frequency: the majority of participants report watching YouTube for 30 minutes to 1 hour each day.

```{r}
#| label: fig-time-distribution
#| fig-cap: "Distribution of participants' YouTube watch time"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4

# Set figure output settings - renders really small

ggplot(subset(youtube, !is.na(Ctime)), aes(x = Ctime)) + 
  geom_bar(color = "black", fill = "grey") +
  facet_grid(~ daily) +
  scale_x_discrete(labels = c("About 15 minutes a day" = "~ 15 min", "About 30 minutes a day" = "~ 30 min", "About 45 minutes a day" = "~ 45 min", "About 1 hour a day" = "~ 1 hr", "About 1 and 1/2 hours a day" = "~ 1.5 hr", "More than 2 hours a day" = "> 2 hr")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Time Spent on YouTube (Self-Reported)",
       y = "Number of Participants",
       title = "Time Spent on YouTube on an Average Day",
       subtitle = "Comparing children who use YouTube every day and children who do not") +
  my_theme()

```

```{r}
#| label: generate-frequency-table

frequency <- round(addmargins(prop.table(table(youtube$Ctime, youtube$daily))), 2)
```

```{r}
#| label: tbl-frequency
#| tbl-cap: "YouTube Screen Time Frequency Table"

kable(frequency, 
      align = c("c", "c", "c"),
      # Setting caption here
      caption = "Proportions of watch times in children who watch YouTube every day vs those who do not") %>% 
  kable_styling(position = "center", latex_options = c("striped"))
```

```{r}
#| label: anova-outputs

# ANOVA: racial regularities score predicted by time spent on YT (categorical self-report)
aov_reg_ctime <- aov(regularity_score ~ Ctime, data = youtube)

# ANOVA: racial regularities score predicted by perceptions of how often kids see outgroup (categorical)
aov_reg_outfreq <- aov(regularity_score ~ outgroup_freq, data = youtube)

# ANOVA: racial regularities score predicted by perceptions of how often kids see outgroup (categorical)
aov_mall_outfreq <- aov(malleable_score ~ outgroup_freq, data = youtube)
```

```{r}
#| label: anova-report-function

# based on apaquarto manuscript example and chat help

# Takes an aov model object as argument
# Returns string of format "$F$(df1, df2) = F-value, $p$ = p-value"

report_anova <- function(model) {

  # make summary object
  m_summary <- summary(model)[[1]]
  # extract values
  df1 <- m_summary$Df[1]
  df2 <- m_summary$Df[2]
  F_value <- m_summary$`F value`[1]
  p_value <- m_summary$`Pr(>F)`[1]
  # Create a string with the results
  report <- paste0("$F$(", 
                   df1,  ", ", 
                   df2, ") = ", 
                   round(F_value, 2), 
                   ", $p$ = ", round(p_value, 3))
  # Return report
  return(report)    
}

```

Preliminary one-way analysis of variance does not indicate that greater amounts of time spent on YouTube (self-report) is associated with a higher racial regularities score (`r report_anova(aov_reg_ctime)`). Similarly, thus far, one-way analysis of variance does not indicate that there is any association between ratings of frequency with which children encounter those of other racial groups and racial regularities scores, (`r report_anova(aov_reg_outfreq)`). Further, one-way analysis of variance does not indicate a significant relationship between the primary racial attitudes outcome measure, group malleability, and frequency of seeing racial outgroups on YouTube, (`r report_anova(aov_mall_outfreq)`).

```{r}
#| label: regression-models

# Regressing malleability against time spent on YouTube (reported by parents)
lm_mt <- lm(malleable_score ~ Ptime, data = youtube)

# Regressing malleability against time spent on YouTube (reported by parents) and whether or not YouTube is used daily
lm_mtd <- lm(malleable_score ~ Ptime + daily, data = youtube)

# Regressing malleability against regularity
lm_mr <- lm(malleable_score ~ regularity_score, data = youtube)
```

```{r}
#| label: regression-report-functions

# Taken from apaquarto demo

# report_regression takes a lm model object as an argument
# Returns string of format: "$F$(df, df.residual) = F-value, $p$ = p-value, $R^2$ = r-squared"

report_regression <- function(model) {

  # Extract coefficients and p-values
  m_glance <- glance(model)
  # Format the p-value
  p_value <- apa_p(m_glance$p.value, add_equals = TRUE)
  # Create a string with the results
  report <- paste0("$F$(", 
                   m_glance$df, 
                   ", ", m_glance$df.residual, 
                   ") = ", round(m_glance$statistic, 2), 
                   ", $p$ = ", p_value)
  # Return report
  return(report)    
}
```

Regressing malleability scores against parent reports of time spent on YouTube, no significant relationship is observed (`r report_regression(lm_mt)`). . This relationship remains unchanged and similarly insignificant when daily use is added as a predictor (`r report_regression(lm_mtd)`). Such relationships, as well as that of regularity score and YouTube usage, may be observed in @fig-malleable and @fig-regularity. Based on pilot sample alone, there is not enough data to conduct analyses with regard to the effect of YouTuber on racial attitudes, but the modal favorite YouTuber reported has thus far been `r mode(youtube$favorite)` (*N* = `r sum(youtube$favorite == 'MrBeast')`). Preliminary relationships between favorite YouTuber (this YouTuber in particular) as well as malleability and regularity scores may be found in @fig-malleable and @fig-regularity respectively.

```{r}
#| label: fig-malleable
#| fig-cap: "Correlations Between Time Spent On YouTube and Group Malleability Beliefs"
#| fig-scap: "Effects of time, daily use, and favorite YouTuber"
#| fig-note: "MrBeast is the modal favorite YouTuber. No other YouTuber has been listed with enough frequency to report."
#| fig-align: "center"
#| fig-width: 5
#| fig-height: 2

ggplot(data = youtube %>% 
             filter(Ptime > 10)) +
  aes(x = Ptime, y = malleable_score) +
  facet_grid(~ daily, scales = "free_x") +
  stat_smooth(method = "lm", color = "#800000") +
  geom_point(aes(shape = youtuber)) +
  scale_shape_manual(values = c("Not MrBeast" = 16, "MrBeast" = 1)) +
  labs(x = "Time Spent on YouTube (minutes)", y = "Malleability Score",
       shape = "Favorite YouTuber") +
  my_theme()
```

```{r}
#| label: fig-regularity
#| fig-cap: "Correlations Between Time Spent On YouTube and Racial Regularities"
#| fig-scap: "Effects of time, daily use, and favorite YouTuber"
#| fig-note: "MrBeast is the modal favorite YouTuber. No other YouTuber has been listed with enough frequency to report."
#| fig-align: "center"
#| fig-width: 5
#| fig-height: 2

ggplot(data = youtube %>% 
             filter(Ptime > 10)) +
  aes(x = Ptime, y = regularity_score) +
  facet_grid(~ daily, scales = "free_x") +
  stat_smooth(method = "lm", color = "#800000") +
  geom_point(aes(shape = youtuber)) +
  scale_shape_manual(values = c("Not MrBeast" = 16, "MrBeast" = 1)) +
  labs(x = "Time Spent on YouTube (minutes)", y = "Regularity Score",
       shape = "Favorite YouTuber") +
  my_theme()
```

Just as there is not yet an observed relationship between time spent on YouTube and racial regularities, so too is there a lack of relationship between regularities scores and malleability scores in the present sample (`r report_regression(lm_mr)`). The preliminary relationship between racial regularity and malleability scores within the full sample may be found in @fig-full. Preliminary effects of daily YouTube usage and participant race may be found in @fig-faceted.

```{r}
#| label: fig-full
#| fig-cap: "Correlation Between Racial Regularities Score and Malleability Score"
#| fig-scap: "Full Sample"
#| fig-note: "N = 18, only pilot data used."
#| fig-align: "center"
#| fig-width: 4
#| fig-height: 3

ggplot(youtube) +
  aes(x = regularity_score, y = malleable_score) +
  stat_smooth(method = "lm", color = "#800000", linewidth = 0.8, fill = "lightgrey") +
  geom_point() +
  labs(x = "Regularities Score", y = "Malleability Score",
       title = "A") +
  my_theme()
```

```{r}
#| label: fig-faceted
#| fig-cap: "Correlation Between Racial Regularities Score and Malleability Score"
#| fig-scap: "Effects of daily use and participant race"
#| fig-note: "Participant race coded as White/Non-White for pilot data only."
#| fig-align: "center"
#| fig-width: 4
#| fig-height: 3

ggplot(youtube) +
  aes(x = regularity_score, y = malleable_score) +
  stat_smooth(method = "lm", color = "black", linewidth = 0.8, fill = "lightgrey") +
  geom_point(aes(color = race)) +
  scale_color_manual(values = c("#155F83", "#C16622")) +
  facet_wrap(~ daily) +
  labs(x = "Regularities Score", y = "Malleability Score",
       color = "Participant Race",
       title = "B") +
  theme(legend.position = "bottom") +
   my_theme()

```

# Discussion {#sec-discussion}

The present study examines the relationship between children's YouTube usage and potential effects on racial attitudes. No significant effects of YouTube usage on group beliefs have been observed thus far; we thus fail to reject the null hypothesis that there is no difference in racial beliefs between children who spend a lot of time on YouTube and those who do not. Similarly, as there is no significant relationship between time spent on YouTube and children's observation of racial regularities, nor between racial regularities and group malleability beliefs, there is no evidence thus far to suggest that observation of racial regularities affects outgroup prejudices as a result of YouTube usage.

\clearpage
# References
:::{refs}
:::