### 2025-03-31 data/preparation.Rmd
Updating document names to reflect full sample data

---
title: "preparation"
author: "Reilly Amera"
date: "2025-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup}
library(tidyverse)

# Read in data
child <- read_csv("child.csv", col_names = TRUE)
parent <- read_csv("parent.csv", col_names = TRUE)
```

```{r cleaning}
# Remove rows that do not contain survey data
parent <- parent[-c(1, 2),]
child <- child[-c(1, 2),]
```

```{r parent cleaning}
# Parent cleaning
parent <- parent %>% 
  # self consent = agree
  filter(Pconsent_self == "Agree") %>% 
  # child consent = agree
  filter(`Pconsent-child` == "Agree") %>% 
  # remove duplicates
  mutate(Q_RelevantIDDuplicate = 
           # Logical operators
           ifelse(is.na(Q_RelevantIDDuplicate), FALSE, TRUE)) %>% 
  filter(Q_RelevantIDDuplicate != TRUE) %>% 
  # filter attn checks
  filter(yt_opinions_4 == "A lot") %>% 
  # get rid of unnecessary columns
  select(-c(StartDate:Q_RelevantIDLastStartDate))
```

```{r parent wrangling}
# Parent reported child age
parent <- parent %>% 
  # Create one birthday column
  unite(bday, `bday#1_1`, `bday#2_1`, sep = " ", remove = FALSE) %>% 
  unite(bday, bday, `bday#3_1`, sep = ", ", remove = FALSE) %>% 
  # Parse birthdays as dates
  mutate(bday = mdy(bday)) %>% 
  # Age column
  mutate(P_age = time_length(interval(bday, Sys.Date()), "years") %>% floor())
```

```{r child cleaning}
child <- child %>% 
  # remove survey testing responses 
  filter(!is.na(transaction_id)) %>% 
  # consent = agree
  filter(Cconsent == "YES") %>% 
  # get rid of unnecessary columns
  select(-c(StartDate:UserLanguage)) %>% 
  # gc = 1 - embedded data; good/complete responses
  filter(gc == 1)
```

```{r merge}
# Merging data
youtube <- inner_join(parent, child, by = join_by(transaction_id))
```

```{r age wrangling}
youtube <- youtube %>% 
  # Do child ages match parent ages?
  mutate(age_match = ifelse(
    P_age == age,
    TRUE,
    FALSE
  ))

# age discrepancies
false_age <- youtube %>% 
  select(age_match, bday, P_age, age, transaction_id) %>% 
  filter(age_match == FALSE) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(difference = abs(P_age - age)) %>% 
  mutate(large_diff = ifelse(
    difference > 1,
    TRUE,
    FALSE
  )) %>% 
  filter(large_diff == TRUE) %>% 
  select(transaction_id)

# write_csv(false_age, "transaction_ids.csv")
```

```{r youtube wrangling}
# Time Scores
```

```{r youtube composites}
### Recoding factor variables as numeric variables to create aggregate scores
youtube <- youtube %>% 
  mutate(
    across(c(`homophily_appear_1`, `homophily_appear_2`, `homophily_appear_3`, `homophily_appear_4`, `homophily_appear_5`, `homophily_behave_1`, `homophily_behave_2`, `homophily_behave_3`, `homophily_behave_4`, `homophily_behave_6`, parasocial_1, parasocial_2, parasocial_3, parasocial_4, parasocial_5),
           ~ recode (.,
                     "Really disagree" = 1,
                     "Disagree" = 2,
                     "I don't know" = 3,
                     "Agree" = 4,
                     "Really agree" = 5
           ))) %>% 
  mutate(
    across(c(malleable_1, malleable_2, malleable_3, malleable_4),
           ~ recode(.,
                    "Really disagree" = 1,
                    "Disagree" = 2,
                    "Neither agree nor disagree" = 3,
                    "Agree" = 4,
                    "Really agree" = 5
           ))) %>%  
  mutate(
    across(c(regularities_1, regularities_2, regularities_3, regularities_4, regularities_5, regularities_6, regularities_7),
           ~ recode(.,
                    "Really disagree" = 1,
                    "Disagree" = 2,
                    "Neither agree nor disagree" = 3,
                    "Agree" = 4,
                    "Really agree" = 5
           ))) %>%  
  select(-c(homophily_behave_5))

### Composite scores
# Racial attitudes score -- total possible score = 20
youtube <- youtube %>% 
  mutate(malleable_score = rowSums(across(malleable_1:malleable_4))) %>% 
  mutate(malleable_score = malleable_score/20)

# Parasocial relationship scores (each block separately, then aggregate)
youtube <- youtube %>% 
  # homophily-appear -- total possible score = 20
  mutate(h_appear_score = rowSums(across(homophily_appear_1:homophily_appear_5))) %>% 
  # homophily-behave -- total possible score = 20
  mutate(h_behave_score = rowSums(across(homophily_behave_1:homophily_behave_6))) %>% 
  # parasocial -- total possible score = 20
  mutate(parasocial_score = rowSums(across(parasocial_1:parasocial_5))) %>% 
  #Composite scores
  mutate(composite = rowSums(across(h_appear_score:parasocial_score))) %>% 
  # Proportion scores
  mutate(h_appear_score = h_appear_score/20) %>%
  mutate(h_behave_score = h_behave_score/20) %>%
  mutate(parasocial_score = parasocial_score/20) %>%
  mutate(composite = composite/60)

# Racial regularities score (behavior) -- total possible score = 35
youtube <- youtube %>% 
  # Reverse coded items
  mutate(across(c(regularities_3, regularities_5, regularities_7), ~ 6 - .)) %>% 
  # Score
  mutate(regularity_score = rowSums(across(regularities_1:regularities_7))) %>% 
  mutate(regularity_score = regularity_score/35)
```

```{r}
# H1: time spent on YouTube and children’s observation of racial regularities covary
```

```{r}
# H2: if children spend more time on YouTube, they will display greater racial stereotyping and prejudice against outgroups
```

```{r}
# H3: the relationship between children’s time spent on YouTube and their exhibited racial stereotyping and prejudice will be mediated by their observations of negative racial regularities
```

```{r}
# H4: the relationship between the amount of time that children spend on YouTube and their racial stereotyping and prejudice will be moderated by whether they have a parasocial relationship with a favorite YouTuber of a race other than their own
```